---
title: "Plots-final-article1"
author: "Clea Abello"
date: '2024-02-21'
output: html_document
---
### 1. INITIALISATION
# 1.1. Loading libraries from packages

```{r}
rm(list=ls())
#-----------------Loading packages-------------------

# List of required packages
pkgs <- c("tidyverse", "here", "cowplot", "sf", "RColorBrewer", "ggpubr",
          "raster", "terra", "osmose", "ncdf4", "fields", "vegan", 
          "heatmaply", "scales", "RColorBrewer", "ggcorrplot", "landscapemetrics", "zenodor")

# Install CRAN packages (excluding special cases)
cran_pkgs <- setdiff(pkgs, c("osmose", "zenodoR"))
install_missing <- setdiff(cran_pkgs, rownames(installed.packages()))
if (length(install_missing)) install.packages(install_missing, dependencies = TRUE)

# Install 'osmose' from drat if needed
if (!"osmose" %in% rownames(installed.packages())) {
  install.packages("osmose", repos = "https://github.com/osmose-model/drat")
}

# Install 'zenodoR' from GitHub if needed
if (!"zenodoR" %in% rownames(installed.packages())) {
  if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
  remotes::install_github("frbcesab/zenodor")
}

# Load all packages
invisible(lapply(pkgs, require, character.only = TRUE))
```

# 1.2. Check data folder is in the root directory of the repository and download from Zenodo if missing
```{r}
# Define settings
data_folder <- "data"
repo_root <- here::here()
data_path <- as.character(file.path(repo_root, data_folder))

# If data folder is missing, use zenodor to download

# Download if data folder doesn't exist
if (!dir.exists(data_path)) {
  message("Data folder not found. Attempting to download from Zenodo or to unzip zenodo folder")

  if(!file.exists("content")) { 
  # Download all files from the Zenodo record
    message("Downloading from Zenodo... Please wait")
  wanted_folder <- zenodor::zen_list_files(record_id = "14039492")[1]
  zenodor::zen_download_files(record_id = "14039492", path = ".", files = wanted_folder)
  file_path <- file.path(repo_root, "content")
  }

  if(file.exists("content")) {
    message("Data folder successfully downloaded")
    unzip(file_path, exdir = repo_root)
    }
    
  } 

if(dir.exists(data_path)){message("Data folder successfully unzipped")} else {stop("Data folder still missing. Check the contents of the Zenodo archive.")}
    
```


# 1.3. Loading osmose config files for fishing mortality and species names

```{r}
# import classification of species by groups
vertical_distrib = read.csv(here("data/vertical_distribution_species.csv")) %>%
  dplyr::select(species_name, group) %>%
  rownames_to_column("species_id") %>%
  mutate(species_id = as.numeric(species_id) - 1)

groups = unique(vertical_distrib$group)
species_id = vertical_distrib$species_id

# Define parameters
Step = 30 # number of MPA coverage increments from 1 to 30%
NbofYears = 60 # averaged number of years from OSMOSE
Nbofreps = 30 # averaged number of replicates from OSMOSE as the model is stochastic
pas = 62 # number of OSMOSE-MED cells corresponding to 1% of the model grid composed of 6229 cells
```

# 1.4. MPA scenarios name and files and fishing redistribution strategies names
```{r}
lit.scenarios.name <- c("Mazor scenario 8",
                    "Mazor scenario 9",
                    "Micheli",
                    "Existing network")

random.scenarios.name <- c("Random 11",
                    "Random 12",
                    "Random 13",
                    "Random 14",
                    "Random 15",
                    "Random 16",
                    "Random 17",
                    "Random 18",
                    "Random 19",
                    "Random 20",
		                "EEZ-conservation 1",
		                "EEZ-conservation 2",
	                  "EEZ-conservation 3",
                    "EEZ-conservation 4",
                    "EEZ-conservation 5",
                    "EEZ-conservation 6",
                    "EEZ-conservation 7",
                    "EEZ-conservation 8",
                    "EEZ-conservation 9",
                    "EEZ-conservation 10")

Fishing_scenario_name = c("Fishing-the-line", "Proportional by GSA", "Uniform by GSA")
Fishing_scenario = c("Fishing-the-line", "GSA-prop", "GSA-uniform")

MPA_scenarios_name <- c("S5-Mazor GFCM",
                    "S6-Mazor SAUP",
                    "S4-Micheli",
                    "S1-Current")

lit.mpa.mask.dir <- c(here("data/MPA_scenarios/Mazor_scenario_8/csv_gpkg_files"),
                       here("data/MPA_scenarios/Mazor_scenario_9/csv_gpkg_files"),
                       here("data/MPA_scenarios/Micheli/csv_gpkg_files"),
                       here("data/MPA_scenarios/Existing_network/csv_gpkg_files"))


random.mpa.mask.dir <- c(here("data/MPA_scenarios/Random/rep11/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep12/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep13/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep14/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep15/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep16/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep17/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep18/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep19/csv_gpkg_files"),
                       here("data/MPA_scenarios/Random/rep20/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep1/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep2/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep3/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep4/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep5/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep6/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep7/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep8/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep9/csv_gpkg_files"),
                       here("data/MPA_scenarios/EEZ-conservation/rep10/csv_gpkg_files"))

```

# 1.5. Load palettes for plots
```{r}
mpa_colors <- c("Existing network" = "gold", 
                "Random 11" = "gray60", 
                "Random 12" = "gray60",
                "Random 13" = "gray60",
                "Random 14" = "gray60",
                "Random 15" = "gray60",
                "Random 16" = "gray60",
                "Random 17" = "gray60",
                "Random 18" = "gray60",
                "Random 19" = "gray60",
                "Random 20" = "gray60",
                "Random" = "gray60",

                "EEZ-conservation 1" = "black", 
                "EEZ-conservation 2" = "black", 
                "EEZ-conservation 3" = "black", 
                "EEZ-conservation 4" = "black", 
                "EEZ-conservation 5" = "black", 
                "EEZ-conservation 6" = "black", 
                "EEZ-conservation 7" = "black", 
                "EEZ-conservation 8" = "black", 
                "EEZ-conservation 9" = "black", 
                "EEZ-conservation 10" = "black", 
                "EEZ-scale conservation" = "black",

                "Micheli" = "darkorange", 
                "Mazor scenario 8" = "darkslateblue", 
                "Mazor scenario 9" = "darkorchid")


```


## PLOTS

# Load all scripts that allow to create all figures in the main text and the supplementary material (figures are created in the following chunks)
```{r}

figures_scripts_folder = here::here("code/create_figures")

lapply(list.files(path = figures_scripts_folder, pattern = "\\.R$", full.names = TRUE), source)

```

# Load functions that allow to upload data needed to create all figures
```{r}
load_multiple_csv <- function(folder_path) {
  file_list <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
  data_list <- lapply(file_list, read.csv, stringsAsFactors = FALSE)
  return(data_list)
}

tbl_fread <- function(output.Folder, pattern){
  files = list.files(output.Folder, pattern, full.names = TRUE) %>% 
    map_df(~fread(.)) %>%
    bind_rows()
}
```


# Figure 1 - changes in total biomass and catch science in the Mediterranean
```{r}

area = "Entire Mediterranean Sea"

# load data
biomass_data <- load_multiple_csv(here("data/total_biomass"))

Biom_fline <- biomass_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") %>%
  filter(mpa == area)

Biom_GSA_prop <- biomass_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area) 

Biom_GSA_uniform <- biomass_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") %>%
  filter(mpa == area) 


catch_data <- load_multiple_csv(here("data/total_catch"))


Catch_fline <- catch_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") %>%
  filter(mpa == area) 

Catch_GSA_prop <- catch_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area)

Catch_GSA_uniform <- catch_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") %>%
  filter(mpa == area)


# FIGURE 
legend_scenarios <- c("Existing network", "Micheli", "Random 11", "Mazor scenario 8", "EEZ-conservation 1", "Mazor scenario 9")

group_levels <- c("Existing network", "Micheli", "Random", "Mazor scenario 8", "EEZ-scale conservation", "Mazor scenario 9")

legend_labels <- c("S1-Current", "S4-Micheli", "S2-Random basin", "S5-Mazor GFCM", "S3-Random EEZ", "S6-Mazor SAUP")

plot.biomass.catch.change.Med(Biom_GSA_prop, Biom_GSA_uniform, Biom_fline, Catch_GSA_prop, Catch_GSA_uniform, Catch_fline, save = TRUE)


# Wilcoxon statistical test - difference between S1 and S2-S3 at 10% coverage
Statistical.difference.scenarios(Catch_GSA_prop, Catch_GSA_uniform, Catch_fline, coverage_level = 10)

```

For main text: variability across 30 replicates for total biomass and total catch
```{r}
#total biomass
Science_informed_B = rbind(Biom_fline, Biom_GSA_prop, Biom_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Mazor scenario 8", "Mazor scenario 9", "Micheli")) %>%
  mutate(rel_biomass_change = (tot.biomass - tot.biomass.ref)/tot.biomass.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_biomass_change),
            sd = sd(rel_biomass_change))

Science_informed_S5_S6_B = rbind(Biom_fline, Biom_GSA_prop, Biom_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Mazor scenario 8", "Mazor scenario 9")) %>%
  mutate(rel_biomass_change = (tot.biomass - tot.biomass.ref)/tot.biomass.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_biomass_change),
            sd = sd(rel_biomass_change))

Science_informed_S4_B = rbind(Biom_fline, Biom_GSA_prop, Biom_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Micheli")) %>%
  mutate(rel_biomass_change = (tot.biomass - tot.biomass.ref)/tot.biomass.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_biomass_change),
            sd = sd(rel_biomass_change))

Random_B = rbind(Biom_fline, Biom_GSA_prop, Biom_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(grepl("EEZ|Random", scenario)) %>%
  mutate(scenario_group = case_when(
    grepl("Random", scenario) ~ "Random",
    TRUE ~ "EEZ-scale conservation"
  )) %>%
  mutate(rel_biomass_change = (tot.biomass - tot.biomass.ref)/tot.biomass.ref * 100) %>%
  group_by(scenario_group, Fishing_scenario) %>%
  summarize(mean = mean(rel_biomass_change),
            sd = sd(rel_biomass_change))

Current_B = rbind(Biom_fline, Biom_GSA_prop, Biom_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Existing network")) %>%
  mutate(rel_biomass_change = (tot.biomass - tot.biomass.ref)/tot.biomass.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_biomass_change),
            sd = sd(rel_biomass_change))

# total catch
Science_informed_Y = rbind(Catch_fline, Catch_GSA_prop, Catch_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Mazor scenario 8", "Mazor scenario 9", "Micheli")) %>%
  mutate(rel_catch_change = (tot.catches - tot.catches.ref)/tot.catches.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_catch_change),
            sd = sd(rel_catch_change))

Science_informed_Y_S5_S6 = rbind(Catch_fline, Catch_GSA_prop, Catch_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Mazor scenario 8", "Mazor scenario 9")) %>%
  mutate(rel_catch_change = (tot.catches - tot.catches.ref)/tot.catches.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_catch_change),
            sd = sd(rel_catch_change))

Science_informed_Y_S4 = rbind(Catch_fline, Catch_GSA_prop, Catch_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Micheli")) %>%
  mutate(rel_catch_change = (tot.catches - tot.catches.ref)/tot.catches.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_catch_change),
            sd = sd(rel_catch_change))

Random_Y = rbind(Catch_fline, Catch_GSA_prop, Catch_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(grepl("EEZ|Random", scenario)) %>%
  mutate(scenario_group = case_when(
    grepl("Random", scenario) ~ "Random",
    TRUE ~ "EEZ-scale conservation"
  )) %>%
  mutate(rel_catch_change = (tot.catches - tot.catches.ref)/tot.catches.ref * 100) %>%
  group_by(scenario_group, Fishing_scenario) %>%
  summarize(mean = mean(rel_catch_change),
            sd = sd(rel_catch_change))

Current_Y = rbind(Catch_fline, Catch_GSA_prop, Catch_GSA_uniform) %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario %in% c("Existing network")) %>%
  mutate(rel_catch_change = (tot.catches - tot.catches.ref)/tot.catches.ref * 100) %>%
  group_by(scenario, Fishing_scenario) %>%
  summarize(mean = mean(rel_catch_change),
            sd = sd(rel_catch_change))
```


Supplementary Figure S5.3 - variability among 30 replicates for total catch and biomass for FPA scenarios S1, S4-S6
```{r}
plot.biomass.catch.change.Med.with.sd(Biom_GSA_prop, Biom_GSA_uniform, Biom_fline, Catch_GSA_prop, Catch_GSA_uniform, Catch_fline, save = TRUE)
```

# Figure 2 - changes in biomass and catch relative to before MPA establishment at 10% for each fishing redistribution strategy according to the distance to the MPA
```{r}

area = "Entire Mediterranean Sea"

# load data
biomass_data <- load_multiple_csv(here("data/total_biomass"))

Biom_fline <- biomass_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") %>%
  filter(mpa == area)

Biom_GSA_prop <- biomass_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area)

Biom_GSA_uniform <- biomass_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") %>%
  filter(mpa == area)


catch_data <- load_multiple_csv(here("data/total_catch"))


Catch_fline <- catch_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") %>%
  filter(mpa == area)

Catch_GSA_prop <- catch_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area)

Catch_GSA_uniform <- catch_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") %>%
  filter(mpa == area)

# Figure for Mazor scenario 8 at 10% coverage
plot.biomass.catch.change.across.MPA(lit.mpa.mask.dir, scen = 1, mpa_scenario = "Mazor_scenario_8", mpa_coverage = 10, save = TRUE)
```

# Figure 3 - Changes in catch and biomass by groups (3A) and changes in small pelagic biomass inside MPAs (3B)
```{r}

# load data
biomass_by_groups_data <- load_multiple_csv(here("data/biomass_by_groups"))

biomass_by_groups <- biomass_by_groups_data[[2]] #GSA prop

catch_by_groups_data <- load_multiple_csv(here("data/catch_by_groups"))

catch_by_groups <- catch_by_groups_data[[2]] #GSA prop

# FIGURE
sp_group_levels <- c("Medium & large pelagic", "Medium & large demersal", "Medium & large benthic", "Small benthic", "Small pelagic", "Crustacean","Cephalopod", "Small demersal")

legend_scenarios <- c("Existing network", "Random 11", "EEZ-conservation 1", "Micheli", "Mazor scenario 8", "Mazor scenario 9")

group_levels <- c("Existing network", "Random", "EEZ-scale conservation", "Micheli", "Mazor scenario 8", "Mazor scenario 9")

legend_labels <- c("S1-Current", "S2-Random basin", "S3-Random EEZ", "S4-Micheli", "S5-Mazor GFCM", "S6-Mazor SAUP")

plot.fig3(biomass_by_groups, catch_by_groups, scen=1, f=2, save = TRUE) #  scen = 1 - Mazor scenario 8, f = 1 - GSA prop



# other scenarios - Supplementary material S6

# Current and literature (S1, S4-S6)
for (scen in 1:4){
  plot.fig.supplementary.S6.other.scenarios.S1.S4.S5.S6(biomass_by_groups_data, catch_by_groups_data, scen=scen, save = TRUE)

}

# Random basin (average of 10) (S2)
plot.fig.supplementary.S6.other.scenarios.S2(biomass_by_groups_data, catch_by_groups_data, save = TRUE)

# Random EEZ (average of 10) (S3)
plot.fig.supplementary.S6.other.scenarios.S3(biomass_by_groups_data, catch_by_groups_data, save = TRUE)


```

Compute SD variability for relative biomass and catch changes by group for main text - FPA scenario S5 and fishing redistribution GSA prop at 10% coverage
```{r}

biomass_by_groups_sd = biomass_by_groups %>%
  # Combine large and medium pelagic/demersal into new groups
  mutate(group = case_when(
    group %in% c("Large-sized pelagic fish", "Medium-sized pelagic fish") ~ "Large & medium pelagic fish",
    group %in% c("Large-sized demersal fish", "Medium-sized demersal fish") ~ "Large & medium demersal fish",
    TRUE ~ group
  )) %>%
  # Sum biomass within the new combined groups before calculating rel change
  group_by(scenario, group, mpa.coverage, mpa, scenario_rep) %>%
  summarize(
    tot.biomass     = sum(tot.biomass,     na.rm = T),
    tot.biomass.ref = sum(tot.biomass.ref, na.rm = T),
    .groups = "drop"
  ) %>%
  # Now calculate relative change per rep
  mutate(rel_change = (tot.biomass - tot.biomass.ref) / tot.biomass.ref * 100) %>%
  # Then summarize across reps
  group_by(scenario, group, mpa.coverage, mpa) %>%
  summarize(
    rel_change_mean = mean(rel_change, na.rm = T),
    rel_change_sd   = sd(rel_change,   na.rm = T),
    .groups = "drop"
  ) %>%
  filter(mpa == "Entire Mediterranean Sea") %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario == "Mazor scenario 8")



catches_by_groups_sd = catch_by_groups %>%
  # Combine large and medium pelagic/demersal into new groups
  mutate(group = case_when(
    group %in% c("Large-sized pelagic fish", "Medium-sized pelagic fish") ~ "Large & medium pelagic fish",
    group %in% c("Large-sized demersal fish", "Medium-sized demersal fish") ~ "Large & medium demersal fish",
    TRUE ~ group
  )) %>%
  # Sum catch within the new combined groups before calculating rel change
  group_by(scenario, group, mpa.coverage, mpa, scenario_rep) %>%
  summarize(
    tot.catches     = sum(tot.catches,     na.rm = T),
    tot.catches.ref = sum(tot.catches.ref, na.rm = T),
    .groups = "drop"
  ) %>%
  # Now calculate relative change per rep
  mutate(rel_change = (tot.catches - tot.catches.ref) / tot.catches.ref * 100) %>%
  # Then summarize across reps
  group_by(scenario, group, mpa.coverage, mpa) %>%
  summarize(
    rel_change_mean = mean(rel_change, na.rm = T),
    rel_change_sd   = sd(rel_change,   na.rm = T),
    .groups = "drop"
  ) %>%
  filter(mpa == "Entire Mediterranean Sea") %>%
  filter(mpa.coverage == 10) %>%
  filter(scenario == "Mazor scenario 8")
```

Barplots for Suppmat with sd for variability between 30 replicate scenarios
```{r}
# load data
biomass_by_groups_data <- load_multiple_csv(here("data/biomass_by_groups"))

biomass_by_groups <- biomass_by_groups_data[[2]] #GSA prop

catch_by_groups_data <- load_multiple_csv(here("data/catch_by_groups"))

catch_by_groups <- catch_by_groups_data[[2]] #GSA prop

# FIGURE
sp_group_levels <- c("Medium & large pelagic", "Medium & large demersal", "Medium & large benthic", "Small benthic", "Small pelagic", "Crustacean","Cephalopod", "Small demersal")

legend_scenarios <- c("Existing network", "Random 11", "EEZ-conservation 1", "Micheli", "Mazor scenario 8", "Mazor scenario 9")

group_levels <- c("Existing network", "Random", "EEZ-scale conservation", "Micheli", "Mazor scenario 8", "Mazor scenario 9")

legend_labels <- c("S1-Current", "S2-Random basin", "S3-Random EEZ", "S4-Micheli", "S5-Mazor GFCM", "S6-Mazor SAUP")


# Current and literature (S1, S4-S6)
for (scen in 1:4){
  plot.fig.supplementary.S6.other.scenarios.S1.S4.S5.S6.with.sd(biomass_by_groups_data, catch_by_groups_data, scen=scen, save = TRUE)

}

# Random basin (average of 10) (S2)
plot.fig.supplementary.S6.other.scenarios.S2.with.sd(biomass_by_groups_data, catch_by_groups_data, save = TRUE)

# Random EEZ (average of 10) (S3)
plot.fig.supplementary.S6.other.scenarios.S3.with.sd(biomass_by_groups_data, catch_by_groups_data, save = TRUE)

```

Fig 3B (small pelagic biomass change) with variability for scenarios S1 and S4-S6 (Supplementary figure S6. 7)
```{r}

plot.small.pelagics.biomass.change.inside.FPAs.with.sd(biomass_by_groups, f = 2)

```

# Figure 4 - Large Fish Indicator (4A) and taxonomic alpha-diversity (4B) inside and outside MPAs
# PROPOSAL - change to relative changes in total Med (same comparable baseline) + add Figure 5 - specifically to show how much diversity is encompassed by the networks (with evenness as well)
```{r}

# load data

# LFI
LFI_data <- load_multiple_csv(here("data/large_fish_indicator"))

LFI_Fline <- LFI_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") 

LFI_GSA_prop <- LFI_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") 

LFI_GSA_uniform <- LFI_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") 


# Diversity

## With sd -- Med
Mean_diversity_Med_GSA_prop_lit =  read.csv(here("data/diversity/Mean_diversity_Med_lit_scenarios_ed_with_sd.csv"))

Mean_diversity_Med_GSA_prop_random_basin =  read.csv(here("data/diversity/Mean_diversity_Med_random_basin_scenarios_ed_with_sd.csv"))

Mean_diversity_Med_GSA_prop_random_EEZ =  read.csv(here("data/diversity/Mean_diversity_Med_random_EEZ_scenarios_ed_with_sd.csv"))

# FIGURE
legend_scenarios <- c("Existing network", "Micheli", "Random 11", "Mazor scenario 8", "EEZ-conservation 1", "Mazor scenario 9")


group_levels <- c("Existing network", "Micheli", "Random", "Mazor scenario 8", "EEZ-scale conservation", "Mazor scenario 9")


legend_labels <- c("S1-Current", "S4-Micheli", "S2-Random basin", "S5-Mazor GFCM", "S3-Random EEZ", "S6-Mazor SAUP")
  

plot.fig4(height_value = 11, width_value = 11, save = TRUE)

```

For main text: replicates for LFI
```{r}
# With replicates for text
LFI_10 = rbind(LFI_GSA_prop, LFI_GSA_uniform, LFI_Fline) %>%
  
  filter(mpa == "Entire Mediterranean Sea") %>%
  rename(LFI = mean_LFI_60_years) %>%
  ungroup() %>%
  group_by(mpa_coverage) %>%
  summarize(mean = mean(LFI),
            sd = sd(LFI))


LFI_all_fishing_scenarios_lit = rbind(LFI_GSA_prop, LFI_GSA_uniform, LFI_Fline) %>%
  filter(scenario %in% c("Micheli", "Mazor scenario 8", "Mazor scenario 9", "Existing network")) %>%
  mutate(x_scale = mpa_coverage * pas/6229 * 100) %>%
  group_by(mpa_coverage, mpa, scenario, Fishing_scenario) %>%
  summarise(mean_LFI_reps = mean(mean_LFI_60_years),
            sd_LFI_reps = sd(mean_LFI_60_years))

LFI_all_fishing_scenarios_Random = rbind(LFI_GSA_prop, LFI_GSA_uniform, LFI_Fline) %>%
  filter(grepl('Random', scenario)) %>%
  mutate(x_scale = mpa_coverage * pas/6229 * 100) %>%
  group_by(mpa_coverage, mpa, scenario, Fishing_scenario) %>%
  summarise(mean_LFI_reps = mean(mean_LFI_60_years)) %>%
  ungroup() %>%
  group_by(mpa_coverage, mpa, Fishing_scenario) %>%
  mutate(high = max(mean_LFI_reps),
         low = min(mean_LFI_reps),
         mean = mean(mean_LFI_reps)) %>%
  mutate(scenario_group = "Random")

LFI_all_fishing_scenarios_EEZ = rbind(LFI_GSA_prop, LFI_GSA_uniform, LFI_Fline) %>%
  filter(grepl('EEZ', scenario)) %>%
  mutate(x_scale = mpa_coverage * pas/6229 * 100) %>%
  group_by(mpa_coverage, mpa, scenario, Fishing_scenario) %>%
  summarise(mean_LFI_reps = mean(mean_LFI_60_years)) %>%
  ungroup() %>%
  group_by(mpa_coverage, mpa, Fishing_scenario) %>%
  mutate(high = max(mean_LFI_reps),
         low = min(mean_LFI_reps),
         mean = mean(mean_LFI_reps)) %>%
  mutate(scenario_group = "EEZ-conservation")  

LFI_10_lit = LFI_all_fishing_scenarios_lit %>%
  filter(mpa == "Entire Mediterranean Sea") %>%
  ungroup() %>%
  dplyr::select(mpa_coverage, scenario, Fishing_scenario, mean_LFI_reps) %>%
  rename(LFI = mean_LFI_reps)


LFI_10_random = LFI_all_fishing_scenarios_Random %>%
  filter(mpa == "Entire Mediterranean Sea") %>%
  ungroup() %>%
  distinct(mean, .keep_all = T) %>%
  dplyr::select(mpa_coverage, scenario_group, Fishing_scenario, mean) %>%
  rename(LFI = mean,
         scenario = scenario_group)

LFI_10_EEZ = LFI_all_fishing_scenarios_EEZ %>%
  filter(mpa == "Entire Mediterranean Sea") %>%
  ungroup() %>%
  distinct(mean, .keep_all = T) %>%
  dplyr::select(mpa_coverage, scenario_group, Fishing_scenario, mean) %>%
  rename(LFI = mean,
         scenario = scenario_group)

LFI_10 = rbind(LFI_10_lit, LFI_10_random, LFI_10_EEZ) %>%
  group_by(mpa_coverage) %>%
  summarize(mean = mean(LFI),
            sd = sd(LFI))

# inside
LFI_10_lit_inside = LFI_all_fishing_scenarios_lit %>%
  filter(mpa == "Inside MPAs") %>%
  ungroup() %>%
  dplyr::select(mpa_coverage, scenario, Fishing_scenario, mean_LFI_reps) %>%
  rename(LFI = mean_LFI_reps)


LFI_10_random_inside = LFI_all_fishing_scenarios_Random %>%
  filter(mpa == "Inside MPAs") %>%
  ungroup() %>%
  distinct(mean, .keep_all = T) %>%
  dplyr::select(mpa_coverage, scenario_group, Fishing_scenario, mean) %>%
  rename(LFI = mean,
         scenario = scenario_group)

LFI_10_EEZ_inside = LFI_all_fishing_scenarios_EEZ %>%
  filter(mpa == "Inside MPAs") %>%
  ungroup() %>%
  distinct(mean, .keep_all = T) %>%
  dplyr::select(mpa_coverage, scenario_group, Fishing_scenario, mean) %>%
  rename(LFI = mean,
         scenario = scenario_group)

LFI_10_inside = rbind(LFI_10_lit_inside, LFI_10_random_inside, LFI_10_EEZ_inside) %>%
  group_by(mpa_coverage) %>%
  summarize(mean = mean(LFI),
            sd = sd(LFI))

```


Supplementary material - LFI inside/outside/Med with sd between replicates (Figure S7.1)
```{r}
# LFI data
LFI_data <- load_multiple_csv(here("data/large_fish_indicator"))

LFI_Fline <- LFI_data[[1]] %>%
  mutate(Fishing_scenario = "Fishing-the-line") 

LFI_GSA_prop <- LFI_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") 

LFI_GSA_uniform <- LFI_data[[3]] %>%
  mutate(Fishing_scenario = "Uniform by GSA") 

# FIGURE
legend_scenarios <- c("Existing network", "Micheli", "Random 11", "Mazor scenario 8", "EEZ-conservation 1", "Mazor scenario 9")


group_levels <- c("Existing network", "Micheli", "Random", "Mazor scenario 8", "EEZ-scale conservation", "Mazor scenario 9")


legend_labels <- c("S1-Current", "S4-Micheli", "S2-Random basin", "S5-Mazor GFCM", "S3-Random EEZ", "S6-Mazor SAUP")
  

plot.fig.suppmat.LFI.inside.outside.Med.with.sd(height_value = 11, width_value = 11, save = TRUE)
```

Supplementary material - inside/outside (Figure S8.3)
```{r}
## With sd -- inside and outside MPAs (Suppmat)
Mean_diversity_inside_GSA_prop_lit =  read.csv(here("data/diversity/Mean_diversity_inside_lit_scenarios_ed_with_sd.csv"))

Mean_diversity_inside_GSA_prop_random_basin =  read.csv(here("data/diversity/Mean_diversity_inside_random_basin_scenarios_ed_with_sd.csv"))

Mean_diversity_inside_GSA_prop_random_EEZ =  read.csv(here("data/diversity/Mean_diversity_inside_random_EEZ_scenarios_ed_with_sd.csv"))

Mean_diversity_outside_GSA_prop_lit =  read.csv(here("data/diversity/Mean_diversity_outside_lit_scenarios_ed_with_sd.csv"))

Mean_diversity_outside_GSA_prop_random_basin =  read.csv(here("data/diversity/Mean_diversity_outside_random_basin_scenarios_ed_with_sd.csv"))

Mean_diversity_outside_GSA_prop_random_EEZ =  read.csv(here("data/diversity/Mean_diversity_outside_random_EEZ_scenarios_ed_with_sd.csv"))

# plot diversity and evenness inside and outside FPAs
plot.fig4.suppmat(height_value = 11, width_value = 11, save = TRUE)
```


# Figure 5 - RDA absolute indicators - total Med 
Fishing redistribution strategy: Proportional by GSA 
FPA coverage: 10% 
Five indicators: biomass, catch, LFI, diversity, evenness
Labels for final figure were added manually for improve readability.
```{r}
# load data
area = "Entire Mediterranean Sea"

biomass_data <- load_multiple_csv(here("data/total_biomass"))

Biom_GSA_prop <- biomass_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area)


catch_data <- load_multiple_csv(here("data/total_catch"))

Catch_GSA_prop <- catch_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") %>%
  filter(mpa == area)


# LFI
LFI_data <- load_multiple_csv(here("data/large_fish_indicator"))

LFI_GSA_prop <- LFI_data[[2]] %>%
  mutate(Fishing_scenario = "Proportional by GSA") 


# Diversity and evenness
Mean_diversity_Med_GSA_prop_lit =  read.csv(here("data/diversity/Mean_diversity_Med_lit_scenarios_ed_with_sd.csv"))

Mean_diversity_Med_GSA_prop_random_basin =  read.csv(here("data/diversity/Mean_diversity_Med_random_basin_scenarios_ed_with_sd.csv"))

Mean_diversity_Med_GSA_prop_random_EEZ =  read.csv(here("data/diversity/Mean_diversity_Med_random_EEZ_scenarios_ed_with_sd.csv"))

# calculate spatial metrics
metrics_calculated = TRUE

if (metrics_calculated == FALSE){
  
  MPA_scenarios = length(c(lit.mpa.mask.dir, random.mpa.mask.dir))
  scenarios = c(lit.scenarios.name, random.scenarios.name)

  # Load contour of Med grid
  Med_grid_sea_contour = st_read(here("data/shapefiles/Med_grid_sea_contour.shp"))
  plot(Med_grid_sea_contour$geometry)
  
  # Calculate average network distance to coast for 10% coverage
  calculate.network.distance.to.coast.fn(Med_grid_sea_contour, MPA_scenarios, scenarios, lit.mpa.mask.dir, random.mpa.mask.dir, mpa_coverage = 10, save = TRUE)
    
  # Calculate metrics from landscapemetrics R package v 2.1.4
  calculate.metrics.landscapemetrics.fn.2(MPA_scenarios, scenarios, lit.mpa.mask.dir, random.mpa.mask.dir, mpa_coverage = 10, save = TRUE)
}

# load spatial metrics
spatial_metrics <- load_multiple_csv(here("data/spatial_metrics"))

Distance_to_coast = spatial_metrics[[1]]
Metrics_mpa_scenarios = spatial_metrics[[2]]


# Capture extent first
tmp <- plot(spe.rda, scaling = 2, type = "n")
rda1_range <- par("usr")[1:2]
rda2_range <- par("usr")[3:4]
dev.off()

# plot RDA
plot.RDA(scaling_type = 1)
plot.RDA(scaling_type = 2)

# plot heatmap showing area of best compromise to maximize biomass and minimize catch losses

spe.rda <- compute.RDA()

# Check the extracted equations printed to console
res <- compute_heatmap(spe.rda, scaling_type = 2, rda1_range, rda2_range)

# Individual plots
plot_rda_heatmap(spe.rda, variable = "biomass", grid_res = 100,
                 save_path = here("figures/RDA_heatmap_biomass.png"))

plot_rda_heatmap(spe.rda, variable = "catch", grid_res = 100,
                 save_path = here("figures/RDA_heatmap_catch.png"))

plot_rda_heatmap(spe.rda, variable = "joint", grid_res = 100,
                 save_path = here("figures/RDA_heatmap_joint.png"))


```


# Figure 5 - Spatial metrics selection
Warning: can take time (~ 20 minutes) to obtain the list of selected variables.
```{r}
Response_vars = create.response.var.matrix(coverage_level = 10)

All_explanatory_var = Metrics_mpa_scenarios %>%
    dplyr::select(metric, value, scenario) %>%
    pivot_wider(names_from = metric, values_from = value) %>%
    left_join(Distance_to_coast, by="scenario") %>%
    dplyr::select(-iji) %>%
    rename("distance_to_coast" = "mean_distance_to_coast") %>%
    column_to_rownames(var="scenario")

All_explanatory_var_standardized = decostand(All_explanatory_var, method = "standardize")

Response_vars <- Response_vars[match(rownames(All_explanatory_var_standardized), rownames(Response_vars)), ]


spe.rda <- rda(Response_vars ~ ., data = All_explanatory_var_standardized)

set.seed(123)  

backward.sel <- ordistep(rda(Response_vars ~ ., data = All_explanatory_var_standardized), # lower model limit (simple!)
                         scope = formula(spe.rda), # upper model limit (the "full" model)
                         direction = "backward",
                         R2scope = TRUE, # can't surpass the "full" model's R2
                         pstep = 1000,
                         trace = TRUE) # change to TRUE to see the selection process!

# Set number of repetitions
n_repeats <- 100
seeds <- sample(1:10000, n_repeats)

# Store selected variables from each iteration
selected_vars_list <- vector("list", n_repeats)

for (i in 1:n_repeats) {
  set.seed(seeds[i])

  # Full model
  full_model <- rda(Response_vars ~ ., data = All_explanatory_var_standardized)

  # Stepwise backward selection
  step_model <- ordistep(
    rda(Response_vars ~ ., data = All_explanatory_var_standardized),
    scope = formula(full_model),
    direction = "backward",
    R2scope = TRUE,
    pstep = 1000,
    trace = FALSE  # Change to TRUE to observe process
  )

  # Extract selected variables
  selected_vars_list[[i]] <- attr(step_model$terms, "term.labels")
}

# Count selection frequencies
selected_counts <- table(unlist(selected_vars_list))
selected_freq <- sort(selected_counts / n_repeats, decreasing = TRUE)

# Print selection frequencies
print(selected_freq)

# Get all unique variables that were selected at least once
selected_vars <- names(selected_freq)

# Print the list
print(selected_vars)
length(selected_vars)

### Variable selection -----------------------------
Selected_explanatory_var <- All_explanatory_var_standardized %>%
    dplyr::select(all_of(selected_vars))

# correlation Heatmap of all variables (Supplementary Figure 8.1)
cor_matrix <- cor(Selected_explanatory_var, use = "complete.obs")
ggcorrplot(cor_matrix,
           type = "full",
           lab = TRUE,
           method = "square",
           colors = c("#6D9EC1", "white", "#E46726"),
           title = "Pearson Correlation Heatmap",
           hc.order = TRUE)

ggsave(here("figures/appendix/Supplementary_Figure_8.1_heatmap_all_vars.png"), dpi = 300, height = 10, width = 10)

rda_final <- rda(Response_vars ~ ., data = Selected_explanatory_var)
RsquareAdj(rda_final)
summary(rda_final)

# Step 1 - Remove highly correlated variables to pladj (correlation > 0.6)
Selected_explanatory_var <- All_explanatory_var_standardized %>%
    dplyr::select(all_of(selected_vars), -total_distance_to_coast, -pd, -gyrate_mn, -para_cv, -mesh, -lpi, -pafrac, -gyrate_sd)

# correlation Heatmap of all variables (Supplementary Figure 8.1)
cor_matrix <- cor(Selected_explanatory_var, use = "complete.obs")
ggcorrplot(cor_matrix,
           type = "full",
           lab = TRUE,
           method = "square",
           colors = c("#6D9EC1", "white", "#E46726"),
           title = "Pearson Correlation Heatmap",
           hc.order = TRUE)


# Step 2 - Remove highly correlated to distance_to_coast
Selected_explanatory_var <- All_explanatory_var_standardized %>%
    dplyr::select(all_of(selected_vars), -total_distance_to_coast, -pd, -gyrate_mn, -para_cv, -mesh, -lpi, -pafrac, -gyrate_sd, -frac_mn, -ndca, -para_mn, -shape_cv, -para_sd, -shape_sd)

# correlation Heatmap of all variables (Supplementary Figure 8.1)
cor_matrix <- cor(Selected_explanatory_var, use = "complete.obs")
ggcorrplot(cor_matrix,
           type = "full",
           lab = TRUE,
           method = "square",
           colors = c("#6D9EC1", "white", "#E46726"),
           title = "Pearson Correlation Heatmap",
           hc.order = TRUE)



# Final selection of variables according to multicollinearity
Final_selected_explanatory_var <- All_explanatory_var_standardized %>%
  dplyr::select("pladj", "np", "shape_mn", "distance_to_coast")

cor_matrix <- cor(Final_selected_explanatory_var, use = "complete.obs")
ggcorrplot(cor_matrix,
           type = "full",
           lab = TRUE,
           method = "square",
           colors = c("#6D9EC1", "white", "#E46726"),
           title = "Pearson Correlation Heatmap",
           hc.order = TRUE)

ggsave(here("figures/appendix/Supplementary_Figure_8.2_heatmap_selected_vars.png"), dpi = 300, height = 10, width = 10)

rda_final <- rda(Response_vars ~ ., data = Final_selected_explanatory_var)
RsquareAdj(rda_final)
summary(rda_final)

plot(rda_final, scaling = 1)
plot(rda_final, scaling = 2)

```

# Figure S2.2 - cumulated ship density per pixel
```{r}
shipping_routes = raster(here("data/SAR_Mediterranean/shipping/shipping.tif")) # WGS84

files <- list.files(here("data/SAR_Mediterranean"), pattern="Rdata", full.names = T)
data_list = lapply(files, load, .GlobalEnv)
sar_2019 <- rbind(S1A_2019_medsea_detections, S1B_2019_medsea_detections)

files.source = list.files(here("code/fishing_effort/process"), full.names = T)
sapply(files.source, source)

sar_all_cleaned = data_clean(sar_2019) # WGS84


# plot
shipping_plot(shipping_routes, sar_all_cleaned, SAR_data_noRFI, save = TRUE)
```

# Figure S5.1 & S5.2 - biomass and catch variogram
```{r}
area = "Entire Mediterranean Sea"

# load data
biomass_data <- load_multiple_csv(here("data/total_biomass"))

Biomass <- biomass_data[[2]] %>% # proportional redistribution strategy
  filter(mpa == area)

catch_data <- load_multiple_csv(here("data/total_catch"))


Catch <- catch_data[[2]] %>%
  filter(mpa == area) 

# fig S5.1

plot.figS5.1(Biomass, Catch, save = TRUE)

# fig S5.2

all_mpa_scenarios = c(lit.scenarios.name, random.scenarios.name)
plot.figS5.2(Biomass, Catch, nreps = 30, save = TRUE)

```
# Figure S5.4 - spatial maps of total biomass and catch
```{r}

worldbounds = st_read(here("data/shapefiles/world_map/world-bounds/world-bounds-a.shp"))
worldbounds_3035 = st_transform(worldbounds, 3035)

mpa.mask.dir = lit.mpa.mask.dir 
# mpa.mask.dir = random.mpa.mask.dir

# only for scenario S5- Mazor-GFCM

plot.fig.suppmat.S5.4(yield_save = TRUE, biomass_save = TRUE, size_save = FALSE, mpa_coverage = 10, mpa.mask.dir)
```


# Figure S6.1 - small pelagics analysis
```{r}

outputDir.nompa = file.path(here("data", "small_pelagics_biomass", "Nompa", "biomass_yield"))

outputDir.mpa.10 = file.path(here("data", "small_pelagics_biomass", "Mazor_scenario_8", "output_mpa_10", "biomass_yield"))

outputDir.mpa.30 = file.path(here("data", "small_pelagics_biomass", "Mazor_scenario_8", "output_mpa_30", "biomass_yield"))

plot.fig.S6.1(outputDir.nompa, outputDir.mpa.10, outputDir.mpa.30, vertical_distrib)
```


# Figure S6.2 - predation pressure
```{r}
outputDir.nompa = file.path(here("data/predation_pressure/Pred_pressure_NoMPA"))
outputDir.mpa = file.path(here("data/predation_pressure/Pred_pressure_Mazor8_10pct_GSAprop"))

plot.fig.S6.2(save = TRUE)
```


# Figure Suppmat S8.1 - alpha-taxonomic divesity inside FPAs before and after their establishment
```{r}
plot.S8.1.diversity.before.after.FPAs(save = TRUE)
```

# Figure S8.2 - alpha-taxonomic divesity, evenness and species richness before FPA establishment
```{r}
plot.fig.suppmat.S8.2.diversity.richness.evenness.maps(save = TRUE)
```

# Figure S9.1 - heatmap RDA
```{r}
plot.heatmap.X.vars()
```


